---
title: "My Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows # alternative is 'columns'
    vertical_layout: scroll # adding scrolling feature alternative is 'fill'
runtime: shiny
---


```{r setup, include=FALSE}

install.packages("tidyverse")
install.packages("dplyr")
install.packages("plotly")
install.packages("gganimate")
install.packages("knitr")
install.packages("gridExtra")
install.packages("data.table")
install.packages("shiny")
install.packages("rsconnect")
install.packages("rmarkdown")
install.packages("bookdown")

```
```{r}
# Loading Libraries

library(tidyr)
library(tidyverse)
library(tidyquant)
library(shiny)
library(shinydashboard)
library(flexdashboard)
library(qcc)
library(dplyr)
library(ggplot2)
library(plotly)
library(gganimate)
library(purrr)
library(data.table)
library(knitr)
library(gridExtra)
library(png)
library(gifski)
library(leaflet)

```

Project Introduction 
=======================================================================

### Presented to Sir Sajid Bashir by Muhammad Siddique Reg. 6237

### ABOUT THE PROJECT (ANALYSING OLYMPIC GAME DATA):


In this project a data set of Olympic Games from 1896 to 2016 will be used.
Attributes Information. This data set contains all the detail of athletes participated, country & city event took place, event year & season etc.


Column {data-width=650} 
-----------------------------------------------------------------------
###	Purpose of the project

```{r}
tags$ol("This set of data will be used to have useful insights and visualizations.")
tags$ol("The Data will be mined and analyzed to answer different questions, like:")
tags$ol("Which country won more medals?")   	
tags$ol("Which player won most medals?")
tags$ol("Comparing performance year wise.")
tags$ol("Visualizing top 10 countries.")
tags$ol("Which city hosted the event most times?")
tags$ol("Medal tally of all / few countries across the years.")
tags$ol("And so many.")

```


Data Exploration
======================================================================

Column {data-width=650} 
-----------------------------------------------------------------------
### TABULAR DATA
```{r}
athletes = read.csv("athlete_events.csv", stringsAsFactors = F)
col_types = cols(
  ID = col_character(),
  Name = col_character(),
  Sex = col_factor(levels = c("M","F")),
  Age =  col_integer(),
  Height = col_double(),
  Weight = col_double(),
  Team = col_character(),
  NOC = col_character(),
  Games = col_character(),
  Year = col_number(),
  Season = col_factor(levels = c("Summer","Winter")),
  City = col_character(),
  Sport = col_character(),
  Event = col_character(),
  Medal = col_factor(levels = c("Gold","Silver","Bronze"))
)

# LOAD DATA MATCHING NOCs (NATIONAL OLYMPIC COMMITTEE) WITH COUNTRIES

NOCs <- read_csv("noc_regions.csv", col_types = cols(
  NOC = col_character(),
  region = col_character()
))

head(athletes)

```
Column {data-width=650} 
-----------------------------------------------------------------------
### Clearing Data by removing NAs from "Medal" Column
```{r}
cleared_data = athletes %>%
  filter(Medal != "<NA>") ## retain all data which is not NA
cleared_data$Year =as.factor(cleared_data$Year)
head(cleared_data)

```
### Medals won by Pakistan
```{r}
medalsbyPak = subset(cleared_data, Team == "Pakistan")
head(medalsbyPak)
```
### Medals won by each country
```{r}
countrywiseMEDALS = cleared_data %>% group_by(Team) %>% 
  summarise(Medals = length(Medal))%>% 
  arrange(desc(Medals))
head(countrywiseMEDALS)
```
### Medals won by each country
```{r}
peryearMedals = cleared_data %>% 
  group_by(Team, Year) %>%
  summarise(Total = n())
head(peryearMedals)

```
Column {data-width=650} 
-----------------------------------------------------------------------
### Medals by selected countries (USA", "Russia", "China", "Germany", "France") per year
```{r}
medalsbyselectedCOUNTRIES = peryearMedals %>% filter(Team %in% c("USA", "Russia", "China", "Germany", "France"))
head(medalsbyselectedCOUNTRIES)
```
### Top-30 Seasonnwise (Summer - Winter) Medals by all countries
```{r}
Seasonnwise = cleared_data %>% group_by(Team, Season) %>%
  summarise(Medals_by_season = length(Medal)) %>%
  arrange (desc(Medals_by_season))
top_n(Seasonnwise,30)
```
### Men Vs Women Performance in the olympics
```{r}
performance = cleared_data %>% group_by(Year, Season, Sex) %>%
  dplyr::summarise(Count=n()) %>%
  arrange(Year)
top_n(performance, 10)
```
Column {data-width=650} 
-----------------------------------------------------------------------
### Men Vs Women Performance in the olympics in Summer 
```{r}
performance_summer = cleared_data %>% group_by(Year, Season, Sex) %>%
  filter(Season == 'Summer') %>%
  dplyr::summarise(Count=n()) %>%
  arrange(Year)
head(performance_summer)
```
### Men Vs Women Performance of top 10 in the olympics 
```{r}
performance_top = cleared_data %>% group_by(Year, Season, Sex) %>%
  top_n(10, Team) %>%
  dplyr::summarise(Count=n()) %>%
  arrange(Year)
head(performance_top)
```
### Most Popular sports by Gender 
```{r}
popularsportsbygender = cleared_data %>% group_by(Sport, Sex) %>%
  filter(Season =='Summer') %>%
  summarize(Count=n()) %>%
  group_by(Sex) %>% top_n(10, Sport)
head(popularsportsbygender)
```
### Medals won by women in 2016
```{r}
f_2016 = cleared_data %>% filter( Year==2016, Sex=="F") %>%
  group_by(NOC, Medal) %>%
  summarize(Count=length(Medal))
head(f_2016)
```
### Medals won by Men in 2016
```{r}
m_2016 = cleared_data %>% filter( Year==2016, Sex=="M") %>%
  group_by(NOC, Medal) %>%
  summarize(Count=length(Medal))
head(m_2016)
```
Column {data-width=100} 
-----------------------------------------------------------------------
### Number of athletes participated
```{r}
athletesnumbers <- cleared_data %>%
  group_by(Year, Sex) %>%
  summarize(count=n())
head(athletesnumbers)
```
### Medals won by women in 2016
```{r}
m_2016 = cleared_data %>% filter( Year==2016, Sex=="F") %>%
  group_by(NOC, Medal) %>%
  summarize(Count=length(Medal))
head(m_2016)
```
### PARTICIPATION OF MALE AND FEMALE ATHLETES OVER TIME, WITHOUT ART COPETITIONS
```{r}
dataOlympics <- athletes %>% 
  filter(Sport != "Art Competitions")
# AFTER 1992, CHANGE THE YEAR OF THE WINTER GAMES TO COINCIDE WITH THE NEXT SUMMER GAMES. THE TERM "YEAR" CURRENTLY REFERS TO THE OLYMPICS TOOK PLACE
original <- c(1994,1998,2002,2006,2010,2014)
new <- c(1996,2000,2004,2008,2012,2016)
for (i in 1:length(original)) {
  dataOlympics$Year <- gsub(original[i], new[i], dataOlympics$Year)
}
dataOlympics$Year <- as.integer(dataOlympics$Year)
# COUNT NUMBER OF ATHLETES BY SEX AND YEAR
countsSex <- dataOlympics %>% 
  group_by(Year, Sex) %>%
  summarize(Athletes = length(unique(ID)))
countsSex$Year <- as.integer(countsSex$Year)
head(countsSex)
```

Basic Charts {data-orientation=columns}
======================================================================

Column {data-height=650}
-----------------------------------------------------------------------

### Top 10 Medals wining countries

```{r}

countrywiseMEDALS %>% top_n(10) %>% ggplot(aes(x =Medals, y= Team))+
  geom_bar(stat = 'identity', colour="white", fill="red")+theme_dark()

```

### Most Popular sports by Gender

```{r}
ggplot(popularsportsbygender, aes(x=Sport, y=Count, group=Sex))+
  geom_col(aes(fill=Sex))+
  theme(axis.text.x = element_text(angle = 90, hjust =1))
```

### Genderwise Athletes participated  

```{r}
 
athletesnumbers %>% ggplot(aes(x=Year, y=count, fill=Sex))+
  geom_col()+
  coord_flip()+
  ggtitle("Genderwise Athletes participated") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.y = element_text(size = 6))
```

Column {data-height=650}
-----------------------------------------------------------------------
### Seasonnwise (Summer - Winter) Medals for China, USA, Russia, France & Germany

```{r}

Seasonnwise_g5 = Seasonnwise %>% filter(Team %in% c("China", "Russia","USA","France","Germany"))
Seasonnwise_g5 %>% ggplot(aes(x = Team, y = Medals_by_season, fill = Season)) +
  geom_bar(stat = "identity", position = position_dodge())+theme_light()
```

### Ploting medals data by selected countries

```{r}

medalsbyselectedCOUNTRIES %>% ggplot(aes(x = Year, y = Total, group = Team))+
  geom_line(aes(color = Team))+geom_point(aes(color = Team)) + 
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

### Plotting Men Vs Women Performance in the olympics

```{r}

performance %>% ggplot(aes(x=Year, y=Count, group=Sex)) +
  geom_line(aes(color=Sex))+
  geom_point(aes(color=Sex))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Column {data-height=350}
-----------------------------------------------------------------------

### Medals won by men in 2016  

```{r}

ggplot(m_2016, aes(x=NOC, y=Count, fill=Medal))+
  geom_col()+
  coord_flip()+
  scale_fill_manual(values=c("Gold1", "Gray70", "Gold4")) + 
  ggtitle("Medal Counts for women at the 2016 Olympics") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.y = element_text(size = 3))

```

### Medals won by women in 2016  

```{r}
 
ggplot(f_2016, aes(x=NOC, y=Count, fill=Medal))+
  geom_col()+
  coord_flip()+
  scale_fill_manual(values=c("Gold1", "Gray70", "Gold4")) + 
  ggtitle("Medal Counts for women at the 2016 Olympics") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.y = element_text(size = 3))
```

More Charts {data-orientation=columns}
=======================================================================

Column {data-height=650}
-----------------------------------------------------------------------
### Has the number of athletes, countries, and events increased or decreased throughout time?

```{r}
numbers <- cleared_data %>%
  group_by(Year, Season) %>%
  summarize(Nations = length(unique(NOC)), Athletes = length(unique(ID)), Events = length(unique(Event))
  )

plotNations <- ggplot(numbers, aes(x=Year, y=Nations, group=Season, color=Season)) +
  geom_point() +
  geom_line() +
  labs(x = " ", y = "Nations", 
       title="Nations, Athletes and Events", 
       subtitle = "Olympic Games from 1896 to 2016")+
  theme(axis.text.x = element_text(size = 5))
plotAthletes <- ggplot(numbers, aes(x=Year, y=Athletes, group=Season, color=Season)) +
  geom_point() +
  geom_line() +
  xlab("")+
  theme(axis.text.x = element_text(size = 5))
plotEvents <- ggplot(numbers, aes(x=Year, y=Events, group=Season, color=Season)) +
  geom_point() +
  geom_line() +
  theme(axis.text.x = element_text(size = 5)) 
grid.arrange(plotNations, plotAthletes, plotEvents, nrow=3)
medalCounts <- cleared_data %>% filter(!is.na(Medal))%>% 
  group_by(NOC, Medal, Event, Games) %>%
  summarize(isMedal=1)
medalCounts <-  medalCounts %>% 
  group_by(NOC, Medal) %>%
  summarize(Count= sum(isMedal))
medalCounts <- left_join(medalCounts, NOCs, by= "NOC" )
medalCounts <- medalCounts %>% 
  mutate (Team = region)
medalCounts <- medalCounts %>% select( Medal, Team, Count)

```

### Top 10 - Nations with the most medals won in history
```{r}

levelsTeam <- medalCounts %>%
  group_by(Team) %>%
  summarize(Total=sum(Count)) %>%
  arrange(desc(Total)) %>%
  select(Team) %>%
  slice(10:1)
medalCounts$Team <- factor(medalCounts$Team, levels=levelsTeam$Team)
medalCounts <- medalCounts %>% filter(Team != "NA")
# PLOT MEDAL COUNTS
ggplot(medalCounts, aes(x=Team, y=Count, fill=Medal)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values=c("gold3","gray64","sienna")) +
  labs(x = "Nations", y = "Count", 
       title="Top 10 - Nations with the most medals won in history", 
       subtitle = "Olympic Games from 1896 to 2016")
```
### show the change of sport event
```{r}
# show the change of sport event
cleared_data<-cleared_data %>%
  group_by(Year,Season) %>% 
  summarise(sports=n_distinct(Sport),Sport_events=n_distinct(Event))%>%
  mutate(Sport=Season, Event=Season)
  
cleared_data %>% ggplot(aes(x=Year,y=Sport_events)) + 
  geom_line(aes(color=Event)) +
  geom_point(aes(color=Event)) +
  geom_bar(aes(x=Year, y=sports*1.5, fill=Sport),stat = "identity") +
  scale_fill_manual(values=c("red","blue")) +
  scale_colour_manual(values = c("red","blue")) +
  facet_grid(Sport~., scale="free_y") +
  labs(x="Year of Olympic Games")

```


Animated Charts {data-orientation=columns}
=======================================================================

Column {data-height=300}
-----------------------------------------------------------------------

### Animated graph showing

```{r}
 
medalCounts <- cleared_data %>% filter(!is.na(Medal))%>% 
  group_by(NOC, Medal, Event, Games, Year) %>%
  summarize(isMedal=1)
medalCounts <-  medalCounts %>% 
  group_by(NOC, Medal, Year) %>%
  summarize(Count= sum(isMedal))
medalCounts <- left_join(medalCounts, NOCs, by= "NOC" )
medalCounts <- medalCounts %>% 
  mutate (Team = region)
medalCounts <- medalCounts %>% select( Medal, Team, Count, Year)

# ORDERING TEAM BY TOTAL MEDAL COUNT
levelsTeam <- medalCounts %>%
  group_by(Team) %>%
  summarize(Total=sum(Count)) %>%
  arrange(desc(Total)) %>%
  select(Team) %>%
  slice(10:1)
medalCounts$Team <- factor(medalCounts$Team, levels=levelsTeam$Team)
medalCounts <- medalCounts %>% filter(Team != "NA")

# ANIMATED PLOT MEDAL COUNT
plotMedalsAnim<- ggplot(medalCounts, aes(x=Team, y=Count, fill=Medal)) +
  labs(x = "Nations", y = "Count", 
       title='Top 10 - Comparison over time, nations with the most medals', 
       subtitle = 'Olympic Games from 1896 to 2016 - Year: {closest_state}')  +
  transition_states(Year)+
  geom_col() +
  coord_flip() +
  scale_fill_manual(values=c("gold3","gray64","sienna"))
animate(plotMedalsAnim,fps=2)
```




